Thesis Title

### 2022-06-08 chapters/chapter_5.tex
Updates from Overleaf

A collection of de novo assembly tools was compiled, including \ac{OLC} and \ac{dBg} assembly algorithms, the latter including both single k-mer and multiple k-mer value approaches, and hybrid assemblers implementing both algorithms, including both genomic and metagenomic assemblers (Supplemental Table S1). Of these, 12 assemblers were selected based on the date of last update and are implemented in LMAS: ABySS \cite{jackman_abyss_2017} (version 2.3.1), BCALM2 \cite{chikhi_compacting_2016} (version 2.2.3), GATB Minia Pipeline \cite{noauthor_gatbgatb-minia-pipeline_2022} (commit hash 9d56f42) , IDBA-UD \cite{peng_idba-ud_2012} (version 1.1.3), MEGAHIT \cite{li_megahit_2015} (version 1.2.9), MetaHipMer2 \cite{georganas_extreme_2018} (version 2.0.0.65-gaad446d-dirty-AddGtest), metaSPAdes \cite{nurk_metaspades_2017} (version 3.15.3), minia \cite{chikhi_space-efficient_2013} (version 3.2.6), SKESA \cite{souvorov_skesa_2018} (version 2.5.0), SPAdes \cite{bankevich_spades_2012} (version 3.15.3), Unicycler \cite{wick_unicycler_2017} (version 0.4.9) and VelvetOptimiser \cite{seemann_velvetoptimiser_2021} (commit hash 092bdee) (Table \ref{tab:ch5_table1}). The execution commands for each assembler are available as Supplemental Material (see Supplemental Material, Short-read de novo assemblers) and in the documentation \cite{noauthor_short-read_nodate}. 
New assemblers can be added with minimal changes to the pipeline so that LMAS can be expanded as novel algorithms are developed. A template is available to facilitate their integration and a step-by-step guide is included in the documentation \cite{noauthor_add_nodate}. The only two requirements for the addition of a new assembler are the execution command for the assembler for paired-end short-read data and a Nextflow-compatible container with the assembler and any dependencies.

@@

Concerning software for assembly quality assessment currently available, the most widely adopted is QUAST \cite{gurevich_quast_2013}, or when dealing with metagenomic data, its extension metaQUAST \cite{mikheenko_metaquast_2016}, which was also adopted by the CAMI challenges [3,5] \cite{sczyrba_critical_2017, meyer_critical_2021} and suggested in the CAMI Tutorial \cite{meyer_tutorial_2021}. Although several features of these tools overlap with LMAS’ quality assessment components, these differ from LMAS in the sense that they are not a single step workflow allowing a traceable and reproducible assembly of mock communities. Unlike QUAST and metaQUAST, whose purpose is to evaluate assemblies, the purpose of LMAS is to allow users to evaluate assembler performance for a given sample of interest. Supplementary Table S2 shows the comparison of the output and computed assembly quality metrics generated by LMAS, QUAST and metaQUAST.

@@

To illustrate the use of LMAS and evaluate the performance of the chosen assemblers we used the eight bacterial genomes and four plasmids of the ZymoBIOMICS Microbial Community Standards as reference. As input we used the raw sequence reads of mock communities with an even and logarithmic distribution of species, from real sequencing runs \cite{nicholls_ultra-deep_2019} and simulated read datasets, with and without error, matching the distribution of species in each sample \cite{gourle_simulating_2019}. Our dataset is composed of samples ENN (in silico generated evenly distributed without error), EMS (in silico generated evenly distributed with Illumina MiSeq error model), ERR2984773 (evenly distributed real Illumina MiSeq sample), LNN (in silico generated logarithmically distributed without error), LHS (in silico generated logarithmically distributed with Illumina HiSeq error model) and ERR2935805 (logarithmically distributed real Illumina HiSeq sample) (see Supplemental Table S3). Detailed information about the generation of the input samples is available as Supplemental Material (see Supplemental Materials, ZymoBIOMICS microbial community standards, Supplemental Table S4). To evaluate the reproducibility of an assembler performance, the LMAS workflow was run three times for all samples using default parameters, and the resulting data was processed for each sample (see Supplemental Materials, Assessment of Assembly Success) Supplementary Table S5 to Table S10 present an overview of the average global performance per assembler for each sample in LMAS.

@@

Of the 12 de novo prokaryotic assemblers included in LMAS, five stand out as having an overall poor performance: ABySS, BCALM2, MetaHipmer2, minia and VelvetOptimiser. Both ABySS and MetaHipmer2 performed inconsistently with differing resource requirements for the same sample in different runs, namely run time and memory allocation (see Supplemental Materials, Resource Requirements Differ Greatly, Supplemental Figure S2).  Moreover, ABySS failed to produce an assembly for sample ERR2984773 for 1 of the runs (see Supplementary Table S7) and for sample LHS in any of the 3 runs in the time limit of 3 days (see Supplementary Table S9), and MetaHipmer2 failed to produce an assembly for samples LNN and LHS in all 3 runs (see Supplementary Tables S8-S9). VelvetOptimiser generated the highest number of inconsistent contigs across the 3 LMAS runs (Figure \ref{fig:chap5_figure3}, Supplementary Table S11), with 1.69\% of the total contigs created present in only 1 or 2 runs. Although not as extreme as VelvetOptimiser, ABySS (0.52\%), minia (0.14\%), GATBMiniaPipeline (0.32\%), MetaHipMer2 (0.11\%) and IDBA-UD (0.08\%) also showed inconsistencies in contig size.

@@

Regarding the quality assessment of the assemblies produced (Figure \ref{fig:chap5_figure4}, Supplementary Table S12), ABySS, BCALM2 and minia are the only single k-mer \ac{dBg} assemblers in the collection and were found to mostly underperform relative to their multiple k-mer \ac{dBg} counterparts, generally resulting in more fragmented assemblies, although there were significant differences in performance across samples. Among multiple k-mer assemblers, VelvetOptimiser frequently produced a very high number of contigs of very small size (over 99\% of the contigs not surpassing the minimum length of 1,000 \ac{bp}) and therefore a low N50 (an average of 29,768 \ac{bp} versus a global average of 84,114 \ac{bp}) (Supplementary tables S5-S10). Additionally, ABySS and VelvetOptimizer produced contigs with a very large number of Ns, with an average of 1,019 and 3,035 uncalled bases per assembly, respectively. MetaHipMer2, although having overall average metrics in the two evenly distributed mock samples (ENN and EMS, Supplementary Tables S5-S6) where it was able to run successfully, it severely underperformed in the real samples (ERR2984773 and ERR2935805, Supplementary Tables S7 and S10). Generally, the performance scores of the assemblers decreased considerably for the real samples in comparison with the simulated ones, either with or without error. High utilisation of the reads in the dataset is observed for most assemblers, with on average at least 90\% of the reads mapping back to the assembly, except for ABySS, BCALM2, MetaHipMer2 and VelvetOptimiser whose values are in the range of 46-79\%. Despite an overall good performance, SPAdes produced the highest number of misassembled contigs, with an average of 98 and a maximum of 572 (sample ERR2935805, Supplementary \ac{Pls}0), in comparison to the global average of 11 misassembled contigs for all assemblers across all samples.

@@

After excluding the poorly performing assemblers, LMAS includes 3 genomic (SKESA, SPAdes and Unicycler) and 4 labelled as metagenomic specific (GATBMiniaPipeline, IDBA-UD, MEGAHIT and metaSPAdes) de novo prokaryotic assemblers, all implementing multiple k-mer \ac{dBg} algorithms. As observed in Figure \ref{fig:chap5_figure5}, Supplementary Table S13 and Supplemental Figure S3,  there were very significant differences between the best and the worst performing assemblers of each type, with this difference being more pronounced for metagenomic assemblers. The best performing assemblers of each type behaved frequently quite similarly, and the differences between them tended to be attenuated after filtering for contigs <1 k\ac{bp}. Still, for the linearly distributed samples (ENN, EMS and ERR2984773), the overall worst performers tended to be metagenomic assemblers. In contrast, for the logarithmically distributed samples (LNN, LHS and ERR2935805) the opposite was observed, with genomic assemblers tending to be the worst-performing (Figure \ref{fig:chap5_figure5}). For the logarithmically distributed samples, the number of basepairs recovered is significantly lower than expected from their composition for both genomic and metagenomic assemblers, particularly after filtering (Supplementary Table S13), as contigs representing the less abundant species are not recovered by either type of assemblers (see Assembler performance is influenced by replicon abundance in the sample). For this dataset, the fact that an assembler is branded as genomic or metagenomic does not translate into better or worse performance in dealing with these complex samples, but rather characteristics of the individual assemblers themselves determine their performance.

@@

For the eight bacterial genomes present in the samples, even in those with an even distribution of the genomes (ENN, EMS and ERR2984773), variations in the assembly metrics were observed (Figure \ref{fig:chap5_figure6}, Supplemental Figures S4-S6, Supplemental Tables S14-S16).  For all samples in the dataset, the genomes are recovered almost completely, with all replicons being >90\% represented in the resulting assemblies. \textit{Lactobacillus fermentum} is the least represented genome (92.2\%-94.9\%). Most replicon sequences are recovered in <100 contigs, except for \textit{Pseudomonas aeruginosa}, \textit{Escherichia coli} and \textit{Salmonella enterica}, and not considering IDBA-UB, which frequently produces a larger number of contigs when compared to other assemblers. The absolute values of other metrics of assembly quality, such as \ac{LSA}, misassembly events or uncalled bases, are also different between bacterial genomes (Supplemental Tables S14-S16). The fact that \textit{S. enterica} is a closely related species to \textit{E. coli}, with high level of genetic similarity (ANIb >0.8, Supplemental Table S23), could have created difficulties for resolving the assemblies in a mixed sample and lead to the lower coverage observed, the higher number of contigs and the increased number of misassembled contigs identified in these species in some samples. However, in the case of the larger number of contigs of \textit{P. aeruginosa}, no related species are present in the sample and these possibly reflect intrinsic properties of the replicon. Similarly, replicon characteristics could be behind the lower breadth of coverage consistently observed in \textit{L. fermentum} assemblies.

@@

Some genomic regions in several replicons are consistently a challenge for all assemblers. As observed in Figure \ref{fig:chap5_figure8}, all genomes present certain regions that fail to assemble for all tools in all runs, even those generating high-quality draft assemblies. Of all seven assemblers considered, only GATBMiniaPipeline, MEGAHIT and IDBA-UD showed inconsistency in the gaps produced over the 3 LMAS runs (Supplemental Table S17), as expected from producing variable sets of contigs. The regions consistently missing for all assemblers in all runs are rich in repetitive elements, such as rRNA and tRNA coding sequences and mobile genetic elements (Supplemental Table S18), with larger gaps corresponding to tandem sets of these elements. This reflects an intrinsic limitation of short-read sequencing since the length of a read pair is not enough to bridge across the repetitive element, preventing the generation of contigs representing these regions. This is something that could be addressed by the use of long-read sequencing technologies. Despite this, some assemblers are able to produce contigs that represent some of these large tandem regions, such as MEGAHIT and SKESA for \textit{E. faecalis}, and IDBA-UD, MEGAHIT and metaSPADES for \textit{L. monocytogenes}, but such performance is not consistent for all reference replicons. For instance, SKESA fails to assemble two large regions of the \textit{S. enterica} genome that all other assemblers successfully cover.

@@

The logarithmically distributed samples (LNN, LHS and ERR2935805) showed greater variation in the assembly success metrics than the evenly distributed samples (Supplementary Table S8-S10), reflecting the difficulty of recovering sequences of the lowest abundant replicons. For the three replicons with an estimated depth of coverage >15x, a similar pattern is observed in logarithmically distributed samples as in evenly distributed samples, albeit with greater dispersion in the number of contigs generated and with a markedly decreased breath of coverage for some assemblers and samples in the logarithmically distributed samples (Figure \ref{fig:chap5_figure6} and Supplementary Figure S7). Almost no contigs >1000 \ac{bp} were retrieved for replicons with an estimated depth of coverage of <2x resulting in a very low breadth of coverage (<1\%) (Supplementary Table S4, Supplementary Table S22). This leads to a severe underrepresentation of the diversity of the community in the generated contigs, particularly of plasmid sequences due to their smaller length and abundance. This happens despite the greater sequencing depth of these samples versus those with an even distribution (>5-fold difference in the number of reads).

@@

We have compiled a collection of de novo assembly tools, including \ac{OLC} and \ac{dBg} assembly algorithms, with both single k-mer and multiple k-mer value approaches, and hybrid assemblers (Supplemental Table S1). The collection includes both genomic and metagenomic assemblers, developed explicitly to handle metagenomic datasets. The dates of the last release correspond to the ones available in the preparation of this manuscript.

@@

For the analysis of LMAS results, the complete ZymoBIOMICS’ reference genomes were annotated with PROKKA \cite{seemann_prokka_2014} (version 1.14.5), using the species-specific database for each reference sequence when available. The number of tRNA, rRNA and mobile element coding genes is available in Supplemental Table S19. Pairwise comparisons among the set of reference replicons were conducted by calculating the Average Nucleotide Identity (ANI)  through BLASTn (version 2.12.0) \cite{camacho_blast_2009} using pyani \cite{pritchard_pyani_2022} (version 0.2.11) \cite{pritchard_genomics_2016}. The results are available in Supplemental Table S23.

@@

A set of simulated samples were generated from the genomes in the ZymoBIOMICS standard through the InSilicoSeq sequence simulator (version 1.5.2) \cite{gourle_simulating_2019}, including both even and logarithmic distribution, with and without Illumina error model. The error model was obtained from each corresponding real sample depending on the distribution and used to generate the mock data with matching characteristics, including read number and abundance of species in the community (Supplemental Table S4).

@@

The processing of the kraken reports was performed through custom python code \cite{noauthor_lmas_2022} where all the percentage of reads that matched for the species in the dataset were saved, as well as the percentage of unclassified reads. For the \textit{Lactobacillus fermentum}, as in the Standard Kraken database no general Species level classification is available, the percentage of reads was calculated as the sum of all reads aligning to one of the \textit{L. fermentum} subspecies. The rest of the reads that were classified as any other species were saved conjunctively as “Other”. Supplemental Table S20 contains the percentage of classified reads for each of the species in the community, as well as “other” and unclassified reads.

@@

The complete set of results for 3 LMAS runs for the raw sequence reads of mock communities with an even and logarithmic distribution of species, from real sequencing runs \cite{nicholls_ultra-deep_2019} and simulated read datasets, with and without error, matching the intended distribution of species in each sample for the eight bacterial genomes and four plasmids of the ZymoBIOMICS Microbial Community Standards as reference is available in Supplemental Table S21 and S22. For the assessment of the assembly success for each sample, the different metrics for all LMAS runs were combined and descriptive statistics, such as the average value, standard deviation, minimum and maximum, were obtained through Python’s Pandas describe function \cite{noauthor_pandas_nodate, noauthor_pandasdataframedescribe_nodate}. Both global and reference based for each assembler, each reference replicon and each sample (ENN - evenly distributed without error model; EMS - evenly distributed with Illumina MiSeq error model; ERR2984773 - real evenly distributed Illumina MiSeq sample, LNN - logarithmically distributed without error model; LHS - logarithmically distributed  with Illumina HiSeq error model; ERR2935805 - real logarithmically distributed Illumina HiSeq sample). For descriptive statistics on several assembler by each assembly type (genomic or metagenomic) and each assembler algorithm (single or multiple k-mer), the use of median was preferred due to its higher robustness against outliers and the high range of the distribution of the results. Plotly \cite{noauthor_plotly_nodate} was used to compute the graphs aggregating the results obtained. The jupyter notebooks \cite{noauthor_project_nodate} with the data processing and all resulting files are available at \cite{noauthor_lmas_2022}.

@@

\begin{landscape}
\input{tables/chapter_5/suptable1}
\end{landscape}

>>>>

A collection of de novo assembly tools was compiled, including \ac{OLC} and \ac{dBg} assembly algorithms, the latter including both single k-mer and multiple k-mer value approaches, and hybrid assemblers implementing both algorithms, including both genomic and metagenomic assemblers (Supplemental Table \ref{tab:ch5_suptable1}). Of these, 12 assemblers were selected based on the date of last update and are implemented in LMAS: ABySS \cite{jackman_abyss_2017} (version 2.3.1), BCALM2 \cite{chikhi_compacting_2016} (version 2.2.3), GATB Minia Pipeline \cite{noauthor_gatbgatb-minia-pipeline_2022} (commit hash 9d56f42) , IDBA-UD \cite{peng_idba-ud_2012} (version 1.1.3), MEGAHIT \cite{li_megahit_2015} (version 1.2.9), MetaHipMer2 \cite{georganas_extreme_2018} (version 2.0.0.65-gaad446d-dirty-AddGtest), metaSPAdes \cite{nurk_metaspades_2017} (version 3.15.3), minia \cite{chikhi_space-efficient_2013} (version 3.2.6), SKESA \cite{souvorov_skesa_2018} (version 2.5.0), SPAdes \cite{bankevich_spades_2012} (version 3.15.3), Unicycler \cite{wick_unicycler_2017} (version 0.4.9) and VelvetOptimiser \cite{seemann_velvetoptimiser_2021} (commit hash 092bdee) (Table \ref{tab:ch5_table1}). The execution commands for each assembler are available as Supplemental Material (see Supplemental Material, Short-read de novo assemblers) and in the documentation \cite{noauthor_short-read_nodate}. 
New assemblers can be added with minimal changes to the pipeline so that LMAS can be expanded as novel algorithms are developed. A template is available to facilitate their integration and a step-by-step guide is included in the documentation \cite{noauthor_add_nodate}. The only two requirements for the addition of a new assembler are the execution command for the assembler for paired-end short-read data and a Nextflow-compatible container with the assembler and any dependencies.

@@

Concerning software for assembly quality assessment currently available, the most widely adopted is QUAST \cite{gurevich_quast_2013}, or when dealing with metagenomic data, its extension metaQUAST \cite{mikheenko_metaquast_2016}, which was also adopted by the CAMI challenges [3,5] \cite{sczyrba_critical_2017, meyer_critical_2021} and suggested in the CAMI Tutorial \cite{meyer_tutorial_2021}. Although several features of these tools overlap with LMAS’ quality assessment components, these differ from LMAS in the sense that they are not a single step workflow allowing a traceable and reproducible assembly of mock communities. Unlike QUAST and metaQUAST, whose purpose is to evaluate assemblies, the purpose of LMAS is to allow users to evaluate assembler performance for a given sample of interest. Supplementary Table \ref{tab:ch5_suptable} shows the comparison of the output and computed assembly quality metrics generated by LMAS, QUAST and metaQUAST.

@@

To illustrate the use of LMAS and evaluate the performance of the chosen assemblers we used the eight bacterial genomes and four plasmids of the ZymoBIOMICS Microbial Community Standards as reference. As input we used the raw sequence reads of mock communities with an even and logarithmic distribution of species, from real sequencing runs \cite{nicholls_ultra-deep_2019} and simulated read datasets, with and without error, matching the distribution of species in each sample \cite{gourle_simulating_2019}. Our dataset is composed of samples ENN (in silico generated evenly distributed without error), EMS (in silico generated evenly distributed with Illumina MiSeq error model), ERR2984773 (evenly distributed real Illumina MiSeq sample), LNN (in silico generated logarithmically distributed without error), LHS (in silico generated logarithmically distributed with Illumina HiSeq error model) and ERR2935805 (logarithmically distributed real Illumina HiSeq sample) (see Supplemental Table \ref{tab:ch5_suptable3}). Detailed information about the generation of the input samples is available as Supplemental Material (see Supplemental Materials, ZymoBIOMICS microbial community standards, Supplemental Table \ref{tab:ch5_suptable4}). To evaluate the reproducibility of an assembler performance, the LMAS workflow was run three times for all samples using default parameters, and the resulting data was processed for each sample (see Supplemental Materials, Assessment of Assembly Success) Supplementary Table \ref{tab:ch5_suptable5} to Table \ref{tab:ch5_suptable10} present an overview of the average global performance per assembler for each sample in LMAS.

@@

Of the 12 de novo prokaryotic assemblers included in LMAS, five stand out as having an overall poor performance: ABySS, BCALM2, MetaHipmer2, minia and VelvetOptimiser. Both ABySS and MetaHipmer2 performed inconsistently with differing resource requirements for the same sample in different runs, namely run time and memory allocation (see Supplemental Materials, Resource Requirements Differ Greatly, Supplemental Figure \ref{fig:chap5_sup_figure_2}).  Moreover, ABySS failed to produce an assembly for sample ERR2984773 for 1 of the runs (see Supplementary Table \ref{tab:ch5_suptable7}) and for sample LHS in any of the 3 runs in the time limit of 3 days (see Supplementary Table \ref{tab:ch5_suptable9}), and MetaHipmer2 failed to produce an assembly for samples LNN and LHS in all 3 runs (see Supplementary Tables \ref{tab:ch5_suptable8}-\ref{tab:ch5_suptable9}). VelvetOptimiser generated the highest number of inconsistent contigs across the 3 LMAS runs (Figure \ref{fig:chap5_figure3}, Supplementary Table \ref{tab:ch5_suptable11}), with 1.69\% of the total contigs created present in only 1 or 2 runs. Although not as extreme as VelvetOptimiser, ABySS (0.52\%), minia (0.14\%), GATBMiniaPipeline (0.32\%), MetaHipMer2 (0.11\%) and IDBA-UD (0.08\%) also showed inconsistencies in contig size.

@@

Regarding the quality assessment of the assemblies produced (Figure \ref{fig:chap5_figure4}, Supplementary Table \ref{tab:ch5_suptable12}), ABySS, BCALM2 and minia are the only single k-mer \ac{dBg} assemblers in the collection and were found to mostly underperform relative to their multiple k-mer \ac{dBg} counterparts, generally resulting in more fragmented assemblies, although there were significant differences in performance across samples. Among multiple k-mer assemblers, VelvetOptimiser frequently produced a very high number of contigs of very small size (over 99\% of the contigs not surpassing the minimum length of 1,000 \ac{bp}) and therefore a low N50 (an average of 29,768 \ac{bp} versus a global average of 84,114 \ac{bp}) (Supplementary tables \ref{tab:ch5_suptable5}-\ref{tab:ch5_suptable10}). Additionally, ABySS and VelvetOptimizer produced contigs with a very large number of Ns, with an average of 1,019 and 3,035 uncalled bases per assembly, respectively. MetaHipMer2, although having overall average metrics in the two evenly distributed mock samples (ENN and EMS, Supplementary Tables \ref{tab:ch5_suptable5}-\ref{tab:ch5_suptable6}) where it was able to run successfully, it severely underperformed in the real samples (ERR2984773 and ERR2935805, Supplementary Tables \ref{tab:ch5_suptable7} and \ref{tab:ch5_suptable10}). Generally, the performance scores of the assemblers decreased considerably for the real samples in comparison with the simulated ones, either with or without error. High utilisation of the reads in the dataset is observed for most assemblers, with on average at least 90\% of the reads mapping back to the assembly, except for ABySS, BCALM2, MetaHipMer2 and VelvetOptimiser whose values are in the range of 46-79\%. Despite an overall good performance, SPAdes produced the highest number of misassembled contigs, with an average of 98 and a maximum of 572 (sample ERR2935805, Supplementary \ac{Pls}), in comparison to the global average of 11 misassembled contigs for all assemblers across all samples.

@@

After excluding the poorly performing assemblers, LMAS includes 3 genomic (SKESA, SPAdes and Unicycler) and 4 labelled as metagenomic specific (GATBMiniaPipeline, IDBA-UD, MEGAHIT and metaSPAdes) de novo prokaryotic assemblers, all implementing multiple k-mer \ac{dBg} algorithms. As observed in Figure \ref{fig:chap5_figure5}, Supplementary Table \ref{tab:ch5_suptable13} and Supplemental Figure \ref{fig:chap5_sup_figure_3},  there were very significant differences between the best and the worst performing assemblers of each type, with this difference being more pronounced for metagenomic assemblers. The best performing assemblers of each type behaved frequently quite similarly, and the differences between them tended to be attenuated after filtering for contigs <1 k\ac{bp}. Still, for the linearly distributed samples (ENN, EMS and ERR2984773), the overall worst performers tended to be metagenomic assemblers. In contrast, for the logarithmically distributed samples (LNN, LHS and ERR2935805) the opposite was observed, with genomic assemblers tending to be the worst-performing (Figure \ref{fig:chap5_figure5}). For the logarithmically distributed samples, the number of basepairs recovered is significantly lower than expected from their composition for both genomic and metagenomic assemblers, particularly after filtering (Supplementary Table \ref{tab:ch5_suptable13}), as contigs representing the less abundant species are not recovered by either type of assemblers (see Assembler performance is influenced by replicon abundance in the sample). For this dataset, the fact that an assembler is branded as genomic or metagenomic does not translate into better or worse performance in dealing with these complex samples, but rather characteristics of the individual assemblers themselves determine their performance.

@@

For the eight bacterial genomes present in the samples, even in those with an even distribution of the genomes (ENN, EMS and ERR2984773), variations in the assembly metrics were observed (Figure \ref{fig:chap5_figure6}, Supplemental Figures \ref{fig:chap5_sup_figure_4}-\ref{fig:chap5_sup_figure_6}, Supplemental Tables \ref{tab:ch5_suptable14}-\ref{tab:ch5_suptable16}).  For all samples in the dataset, the genomes are recovered almost completely, with all replicons being >90\% represented in the resulting assemblies. \textit{Lactobacillus fermentum} is the least represented genome (92.2\%-94.9\%). Most replicon sequences are recovered in <100 contigs, except for \textit{Pseudomonas aeruginosa}, \textit{Escherichia coli} and \textit{Salmonella enterica}, and not considering IDBA-UB, which frequently produces a larger number of contigs when compared to other assemblers. The absolute values of other metrics of assembly quality, such as \ac{LSA}, misassembly events or uncalled bases, are also different between bacterial genomes (Supplemental Tables \ref{tab:ch5_suptable14}-\ref{tab:ch5_suptable16}). The fact that \textit{S. enterica} is a closely related species to \textit{E. coli}, with high level of genetic similarity (ANIb >0.8, Supplemental Table \ref{tab:ch5_suptable23}), could have created difficulties for resolving the assemblies in a mixed sample and lead to the lower coverage observed, the higher number of contigs and the increased number of misassembled contigs identified in these species in some samples. However, in the case of the larger number of contigs of \textit{P. aeruginosa}, no related species are present in the sample and these possibly reflect intrinsic properties of the replicon. Similarly, replicon characteristics could be behind the lower breadth of coverage consistently observed in \textit{L. fermentum} assemblies.

@@

Some genomic regions in several replicons are consistently a challenge for all assemblers. As observed in Figure \ref{fig:chap5_figure8}, all genomes present certain regions that fail to assemble for all tools in all runs, even those generating high-quality draft assemblies. Of all seven assemblers considered, only GATBMiniaPipeline, MEGAHIT and IDBA-UD showed inconsistency in the gaps produced over the 3 LMAS runs (Supplemental Table \ref{tab:ch5_suptable17}), as expected from producing variable sets of contigs. The regions consistently missing for all assemblers in all runs are rich in repetitive elements, such as rRNA and tRNA coding sequences and mobile genetic elements (Supplemental Table \ref{tab:ch5_suptable18}), with larger gaps corresponding to tandem sets of these elements. This reflects an intrinsic limitation of short-read sequencing since the length of a read pair is not enough to bridge across the repetitive element, preventing the generation of contigs representing these regions. This is something that could be addressed by the use of long-read sequencing technologies. Despite this, some assemblers are able to produce contigs that represent some of these large tandem regions, such as MEGAHIT and SKESA for \textit{E. faecalis}, and IDBA-UD, MEGAHIT and metaSPADES for \textit{L. monocytogenes}, but such performance is not consistent for all reference replicons. For instance, SKESA fails to assemble two large regions of the \textit{S. enterica} genome that all other assemblers successfully cover.

@@

The logarithmically distributed samples (LNN, LHS and ERR2935805) showed greater variation in the assembly success metrics than the evenly distributed samples (Supplementary Table \ref{tab:ch5_suptable8}-\ref{tab:ch5_suptable10}), reflecting the difficulty of recovering sequences of the lowest abundant replicons. For the three replicons with an estimated depth of coverage >15x, a similar pattern is observed in logarithmically distributed samples as in evenly distributed samples, albeit with greater dispersion in the number of contigs generated and with a markedly decreased breath of coverage for some assemblers and samples in the logarithmically distributed samples (Figure \ref{fig:chap5_figure6} and Supplementary Figure \ref{fig:chap5_sup_figure_7}). Almost no contigs >1000 \ac{bp} were retrieved for replicons with an estimated depth of coverage of <2x resulting in a very low breadth of coverage (<1\%) (Supplementary Table \ref{tab:ch5_suptable4}, Supplementary Table \ref{tab:ch5_suptable22}). This leads to a severe underrepresentation of the diversity of the community in the generated contigs, particularly of plasmid sequences due to their smaller length and abundance. This happens despite the greater sequencing depth of these samples versus those with an even distribution (>5-fold difference in the number of reads).

@@

We have compiled a collection of de novo assembly tools, including \ac{OLC} and \ac{dBg} assembly algorithms, with both single k-mer and multiple k-mer value approaches, and hybrid assemblers (Supplemental Table \ref{tab:ch5_suptable1}). The collection includes both genomic and metagenomic assemblers, developed explicitly to handle metagenomic datasets. The dates of the last release correspond to the ones available in the preparation of this manuscript.

@@

For the analysis of LMAS results, the complete ZymoBIOMICS’ reference genomes were annotated with PROKKA \cite{seemann_prokka_2014} (version 1.14.5), using the species-specific database for each reference sequence when available. The number of tRNA, rRNA and mobile element coding genes is available in Supplemental Table \ref{tab:ch5_suptable19}. Pairwise comparisons among the set of reference replicons were conducted by calculating the Average Nucleotide Identity (ANI)  through BLASTn (version 2.12.0) \cite{camacho_blast_2009} using pyani \cite{pritchard_pyani_2022} (version 0.2.11) \cite{pritchard_genomics_2016}. The results are available in Supplemental Table \ref{tab:ch5_suptable23}.

@@

A set of simulated samples were generated from the genomes in the ZymoBIOMICS standard through the InSilicoSeq sequence simulator (version 1.5.2) \cite{gourle_simulating_2019}, including both even and logarithmic distribution, with and without Illumina error model. The error model was obtained from each corresponding real sample depending on the distribution and used to generate the mock data with matching characteristics, including read number and abundance of species in the community (Supplemental Table \ref{tab:ch5_suptable4}).

@@

The processing of the kraken reports was performed through custom python code \cite{noauthor_lmas_2022} where all the percentage of reads that matched for the species in the dataset were saved, as well as the percentage of unclassified reads. For the \textit{Lactobacillus fermentum}, as in the Standard Kraken database no general Species level classification is available, the percentage of reads was calculated as the sum of all reads aligning to one of the \textit{L. fermentum} subspecies. The rest of the reads that were classified as any other species were saved conjunctively as “Other”. Supplemental Table \ref{tab:ch5_suptable20} contains the percentage of classified reads for each of the species in the community, as well as “other” and unclassified reads.

@@

The complete set of results for 3 LMAS runs for the raw sequence reads of mock communities with an even and logarithmic distribution of species, from real sequencing runs \cite{nicholls_ultra-deep_2019} and simulated read datasets, with and without error, matching the intended distribution of species in each sample for the eight bacterial genomes and four plasmids of the ZymoBIOMICS Microbial Community Standards as reference is available in Supplemental Table \ref{tab:ch5_suptable21} and \ref{tab:ch5_suptable22}. For the assessment of the assembly success for each sample, the different metrics for all LMAS runs were combined and descriptive statistics, such as the average value, standard deviation, minimum and maximum, were obtained through Python’s Pandas describe function \cite{noauthor_pandas_nodate, noauthor_pandasdataframedescribe_nodate}. Both global and reference based for each assembler, each reference replicon and each sample (ENN - evenly distributed without error model; EMS - evenly distributed with Illumina MiSeq error model; ERR2984773 - real evenly distributed Illumina MiSeq sample, LNN - logarithmically distributed without error model; LHS - logarithmically distributed  with Illumina HiSeq error model; ERR2935805 - real logarithmically distributed Illumina HiSeq sample). For descriptive statistics on several assembler by each assembly type (genomic or metagenomic) and each assembler algorithm (single or multiple k-mer), the use of median was preferred due to its higher robustness against outliers and the high range of the distribution of the results. Plotly \cite{noauthor_plotly_nodate} was used to compute the graphs aggregating the results obtained. The jupyter notebooks \cite{noauthor_project_nodate} with the data processing and all resulting files are available at \cite{noauthor_lmas_2022}.

@@

\newpage

\input{tables/chapter_5/suptable1}
\input{tables/chapter_5/suptable2}
\input{tables/chapter_5/suptable3}
\input{tables/chapter_5/suptable4}
\input{tables/chapter_5/suptable5}
\input{tables/chapter_5/suptable6}
\input{tables/chapter_5/suptable7}
\input{tables/chapter_5/suptable8}
\input{tables/chapter_5/suptable9}
\input{tables/chapter_5/suptable10}
\input{tables/chapter_5/suptable11}
\input{tables/chapter_5/suptable12}
\input{tables/chapter_5/suptable13}
\input{tables/chapter_5/suptable14}
